Lyman-$\alpha$ photons through rotating outflows

### 2018-03-29 paper/paper.tex
edit

\subsection{Qualitative Trends}
\label{sec:qualitative}
Figure \ref{fig:doppler_shift} summarizes the most important trends of
our simulations. 
The six panels correspond to $\tau=10^6$ and a viewing angle of
$\theta =90^{\circ}$, that is perpendicular to the rotation axis of the
galaxy. 
In every panel the thin black line corresponds to the pure outflow
solution (i.e. without rotation). 
From top to bottom we see the effect of increasing the outflow
velocity, which is the expected increasing asymmetry towards the red
peak.

@@

Finally, the thick gray line correspods to the pure outflow solution
with the Doppler boost added to model rotation's influence.
In all cases the Doppler boost does a good job at capturing the broad
morphological features introduced by rotation: the angle dependence,
the broadening and the intensity increase at the line's center.

In Figure \ref{fig:doppler_shift_theta_0} we show the same results as
in Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
90^{\circ}$, that is parallel to the rotation axis. 
In this case we confirme the result presented by \ref{Garavito14}
showed, namely that pure rotation introduces a strong dependence with
viewing angle, a trend that we find also holds for rotation mixed with
ouflows.

@@

After finding the quantitative influence of the different parameters we
move onto a qualilitative study.
To do this we summarize the line morphology by four different
characteristics: standard deviation (STD), skewness (SKW), bimodality
(BI) and valley/peak ration.
These quantities are defined by the following equations \citep{kokoska1999}.:

@@

\begin{figure*}
\begin{center}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
\end{center}
\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative
  Transfer and the lines represent the Doppler Shift.
  \label{fig:standard_deviation}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
\end{center}
\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
    Transfer and the lines represent the Doppler Shift.
    \label{fig:skewness}}
\end{figure*}

@@

In this appendix we show \lya spectra with fixed \vout (Fig. \ref{fig:varying_rotation_small})
and fixed \vrot (Fig. \ref{fig:varying_outflow_small}). For a larger set of spectra with the
same fixed \vout and \vrot refer to Fig. \ref{fig:varying_rotation} and Fig. \ref{fig:varying_outflow}
respectively.

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation_small}
	\end{center}
	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
		\label{fig:varying_rotation_small}}
\end{figure*}

\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow_small}
	\end{center}
	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
		\label{fig:varying_outflow_small}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vout=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100,
% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100,
% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100,
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation}
	\end{center}
	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
		\label{fig:varying_rotation}}
\end{figure*}

% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
% cada panel tiene una curva para theta=0, 45, 90.
% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75,
\begin{figure*}
	\begin{center}
		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow}
	\end{center}
	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
		\label{fig:varying_outflow}}
\end{figure*}

% MUSE
\begin{figure*}
	\begin{center}
		\includegraphics[height=0.3\textheight]{./figures/appendix/muse_vs_simulations.pdf}
		\includegraphics[height=0.3\textheight]{./figures/appendix/museall_vs_muserange.pdf}
		\includegraphics[height=0.3\textheight]{./figures/appendix/muserange_vs_simulations.pdf}
	\end{center}
	\caption{\textbf{Caption:} Caption.
		\label{fig:varying_outflow}}
\end{figure*}

@@

\cite{DjorgovskiThompson}, \cite{Rhoads00},
\cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
\cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12},
\cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013},
\cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
\cite{Fumagalli2015}.

\section{A new LAE model}

(\cite{Adams72}, \cite{Harrington73}, \cit
e{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12},
\cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91},
\cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12},
\cite{Yajima12}).\\

\subsection{Galaxy's Viewing Angle}

We illustrate this in
Fig. \ref{fig:model}. \\

\begin{figure}[h!]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{./figures/model}
  \end{center}
	\caption{\textbf{Model:} Spherical LAE with tangential and radial velocities due to rotation and outflows, respectively. The galaxy cut is in the $y-z$ plane perspective. The observer is located at an specific viewing angle of the sphere. Only photons with a direction that enters in this range of vision are taken into account to build the observed spectrum.
		\label{fig:model}}
\end{figure}

\section{Results}

\subsection{Resulting simulated spectra}
In order to define the ranges of tauh, vrot and vout, it is necessary
to refer to observational constraints. The common values for typical
LAEs that were set as parameters are in Tab. \ref{tab:values}. We run
CLARA's modified version for all the permutations of these 3
parameters. \\

The behavior of the resulting sets of spectra can be summarized by
Fig. \ref{fig:summary}. We concluded from the simulations a clear
creation of two asymmetric peaks around $V=0$ \kms with the tallest
peak always redshifted. We detected a strong dependence on the outflow
velocity that induces the peak asymmetry. We also detected that the
rotation velocity broadens the line horizontally. Both these effects
have been previously reported separately in the literature, so these
results can valid their superposition to some extend. \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.48\textwidth]{./figures/summary}
	\end{center}
	\caption{\textbf{4 different \lya profiles:} With tauh$=10^5$ and $\theta \simeq 90^\circ$. The rotational velocity vrot increases to the right and the outflows velocity vout increases downwards. The intensity is in arbitrary units.
		\label{fig:summary}}
\end{figure}

\subsection{Influence of the viewing angle $\theta$}
We take now into account the viewing angle of the galaxy to build the observed spectra. For all of the physical parameters' combinations, the effect of $\theta$ in the \lya line is always the same.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/angles}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$.} For tauh$=10^5$, vrot$=50$ \kms and vout$=20$ \kms. For tauh$=10^6$, vrot$=100$ \kms and vout$=5$ \kms. For tauh$=10^7$, vrot$=100$ \kms and vout$=15$ \kms. The intensity is in arbitrary units.
		\label{fig:angles}}
\end{figure}

From Fig. \ref{fig:angles} it is clear that the intensity of the valley between the two peaks increases along with $\theta$. This causes an intensity decrease in the rest of the frequencies, thus a broadening of the line. The asymmetry also changes with the viewing angle.\\

\subsection{Morphology of \lya line}

To summarize, the influence of the 4 parameters on the \lya morphology is the following:

\begin{itemize}
	\item tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
	\item vout decreases the right peak's intensity. Higher vout makes the left peak smaller until it merges with the right one. \\
	\item vrot broadens the line and decreases the maximum intensity. Higher vrot implies a flatter spectrum. This effect has not been deeply studied in literature. Only Garavito et al. \cite{Garavito14} has simulated its effect. Our results are consistent with their conclusions.
	\item $\theta$ increases the central valley of the line. Higher $\theta$ makes more emission at \lya natural frequency. \\
\end{itemize}

\subsection{Doppler shift by rotation}

The rotation effect induces a Doppler shift in frequency and displaces the only-outflows spectra. As seen in Fig. \ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location. Then it weights the lines and they all merge into the resulting red \lya line.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.3\textwidth]{./figures/rotation_doppler_outflow}
	\end{center}
	\caption{\textbf{Rotations induces a Doppler shift of the only-outflows spectra:} Each of the black lines is then weighted to obtain the red line.
		\label{fig:rotation_doppler_outflow}}
\end{figure}

PONER GRAFICA REAL...\\

\section{Observational Implications}

Among the observational implications that this paper has over observations we select three principal ones. First, the rotational effects of the galaxy should be clearly detected and characterized by MUSE's high resolution data. The Doppler redshift is already visible in other observations, such as the one seen in Fig. 7 of Prescott et al. \cite{Prescott14}.\\

Second, the central ($V=0$) emission of the spectra seen in the results is a consequence of the viewing angle of the galaxy and can be controlled by it. This means that the reason is not necessary that radiation is escaping without scattering, as several authors have suggested (CITAS...).\\

Finally, all the spectra produced from this new LAE model is roughly consistent with observations. The consideration of adding the new rotation parameter to the standard only-outflows model found in literature is reasonable and very powerful. vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\

\section{Conclusions}

In this paper, the objective was to analyze and measure the influence
of galaxy rotation and outflows on the \lya line. The motivation for
this is to be able to obtain physical information of a LAE by just
looking at its \lya profile. In order to accomplish this objective, we
propose a new model of a LAE consisting of a sphere of Hydrogen atoms
that expands radially and rotates as a solid body. A modified version
of the program CLARA \cite{CLARA} is used to set the conditions and
emulate the radiative transfer process inside the galaxy. \\

The conclusions obtained from this work are:

\begin{itemize}
	\item The outgoing spectra depend on the angle an external observer is viewing the galaxy from. The closer it is to the equator of the galaxy, the higher the central valley of the frequency distribution.

	\item The effects of vrot, vout and tauh are consistent with the different authors that have used them. vrot broadens the \lya line. vout increases the peaks asymmetry and tauh induces a redshift around the zero velocity.

	\item Rotations induces a Doppler shift in frequency of the only-outflows spectra.

	\item The final spectra obtained are roughly consistent with LAEs observations. \\

\end{itemize}

\subsection{Future work}

For future work regarding the model, we would like to implement a differential rotation for the LAE instead of solid body. Regarding result analysis, we would like to compare against observations in two ways. Firstly by using the kinematic observation obtained from newly observed LAEs. We could do a consistency check between their real spectra and the simulated one that our model would produce with the respective parameters. Secondly we would like to take a specific observed spectrum and fit it using MCMC in order to predict ranges for its parameters. \\

\subsection{Reproducibility}

All of this work is available online and free to use to anyone. The data, source code and instructions to replicate this paper's results are in the GitHub repository:  \texttt{github.com/astroandes/CLARA\_RotationOutflows}. \\

\section*{Acknowledgments}
The authors would like to acknowledge the high performance computational cluster (HPC) hosted by Universidad de los Andes, where we ran the simulations for CLARA. \\

We would also like to acknowledge the participants of the meeting \textit{Escape of Lyman radiation from galactic labyrinths} at Crete, Greece, for their valuable comments on this work, specially to Max Gronke and Christian Herenz. \\

%------------------------REFERENCES----------------------------

>>>>

\subsection{Qualitative Trends}
\label{sec:qualitative}
Figure \ref{fig:doppler_shift} summarizes the most important trends.
The six panels correspond to $\tau=10^6$ and a viewing angle of
$\theta =90^{\circ}$, that is, perpendicular to the rotation axis of the
galaxy. 
In every panel the thin black line corresponds to the pure outflow
solution, i.e. without rotation. 
From top to bottom we see the effect of increasing the outflow
velocity, which is the expected increasing asymmetry towards the red
peak.

@@

Finally, the thick gray line correspods to the pure outflow solution
with the Doppler boost added to model rotation's influence.
At $\tauh=10^6$ the Doppler boost does a good job at capturing the broad
morphological features introduced by rotation: the angle dependence,
the broadening and the intensity increase at the line's center.

In Figure \ref{fig:doppler_shift_theta_0} we show the same results as
in Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
90^{\circ}$, that is parallel to the rotation axis. 
In this case we confirmethe result presented by \cite{Garavito14},
namely that pure rotation introduces a strong dependence with 
viewing angle, a trend that we find also holds for rotation mixed with
ouflows.

The quality of the results from the Doppler boos improves for higher
\tauh values. 
In the Appendix we show the same plots as Figures
\ref{fig:doppler_shift} and \ref{fig:doppler_shift}, there it is
evident that for $\tauh=10^5$ the results are not as good as they are
for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler boos
provides a remarkable good approximation.

@@

After finding the quantitative influence of the different parameters we
move onto a qualitative study.
To do this we summarize the line morphology by four different
characteristics: standard deviation (STD), skewness (SKW), bimodality
(BI) and valley/peak ration.
These quantities are defined by the following equations \citep{kokoska1999}.:

@@

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
\end{center}
\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative
  Transfer and the lines represent the Doppler Shift.
  \label{fig:standard_deviation}}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
\end{center}
\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
    Transfer and the lines represent the Doppler Shift.
    \label{fig:skewness}}
\end{figure*}

@@

@@
